Synthesizing Quotient Lenses

### 2018-03-14 papers/QuotientLenses/main.tex
Redid Intro

\lstset{ language=Caml, basicstyle=\linespread{0.8}\upshape\sffamily,
keywordstyle=\upshape\sffamily\color{dkblue}, keepspaces=true,
framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
commentstyle=\itshape\rmfamily,
emph={synth,project,perm,squash,normalize,using,ins,del,lens,let},
emphstyle=\upshape\sffamily\color{dkred}, 
columns=fullflexible,
escapechar=\#,
mathescape, 
xleftmargin=1.5em,
% BCP: I find this distracting:
% stringstyle=\sffamily\color{dkred},
}
\makeatletter
     \let\lst@oldvisiblespace\lst@visiblespace
     \def\lst@visiblespace{\,\lst@oldvisiblespace\,}
\makeatother

@@

Building on recent work that showed how to synthesize bijective lenses from high
level specifications, we offer three innovations to synthesize quotient lenses.
First, we introduce {\em quotient regular expressions} (QREs), that easily mark
aspects of data formats that are inessential and hence give a high-level
specification of regular languages with an equivalence relation defined on
them. Next, we introduce {\em QRE lenses}, i.e., lenses mapping between QREs.
Our key technical result is a proof that every QRE lens can be transformed into
one of the form $c_1 ; \ell ; c_2^{-1}$, where $c_1$ and $c_2^{-1}$ are {\em
canonizers}, $\ell$ is a bijective transformation, and $;$ is
composition\iflastminute\bcp{$c_2^{-1}$ is ugly, but I think we need the
superscript for accuracy.  Perhaps calling them $c$ and $d$ instead of $c_1$
and $c_2$ would be better?}\fi.  Intuitively, the canonizers normalize data by
picking a canonical representative of the relevant equivalence relations. This
theorem means we can push all normalization to the outer edges of a
bidirectional transformation, leaving in the middle a pure bijective
transformation. Finally, we leverage this normalization theorem to {\em
synthesize} QRE lenses from a pair of QREs and example input-output pairs,
reusing earlier work on synthesizing plain bijective lenses. We have
implemented QREs and QRE lens synthesis in the bidirectional programming
language Boomerang. We evaluate the effectiveness of our approach by using it
to synthesize QRE lenses between various real-world data formats in the
Optician benchmark suite.

>>>>

\lstset{ language=Caml, basicstyle=\linespread{0.8}\upshape\sffamily,
keywordstyle=\upshape\sffamily\color{dkblue}, keepspaces=true,
framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
commentstyle=\itshape\rmfamily,
emph={synth,project,perm,squash,normalize,using,ins,del,lens,let},
emphstyle=\upshape\sffamily\color{dkred}, 
columns=fullflexible,
escapechar=\#,
mathescape, 
xleftmargin=1.5em,
% BCP: I find this distracting:
% stringstyle=\sffamily\color{dkred},
}

@@

We offer three interconnected innovations to simplify the task of programming
quotient lenses.  
First, we introduce {\em quotient regular expressions} (QREs), to
easily mark aspects of data formats that are
inessential.  A QRE is an augmented regular expression that
simultaneously defines a regular language and an equivalence
relation on this language.  Variations within the same
equivalence class are deemed inessential. 
Next, we introduce {\em QRE lenses}, i.e., lenses mapping
between QREs.
Our key technical result is a proof that every QRE lens can be
transformed into one of the form $c_1 ; \ell ;
c_2^{-1}$, where $c_1$ and $c_2^{-1}$ are {\em canonizers}, $\ell$ is a bijective
transformation, and $;$ is composition\iflastminute\bcp{$c_2^{-1}$ is ugly, but I think
  we need the superscript for accuracy.  Perhaps calling them $c$ and $d$
  instead of $c_1$ and $c_2$ would be better?}\fi.  Intuitively,
the canonizers normalize data by picking a
canonical representative of the relevant equivalence relations. 
This theorem means we can push all normalization to the outer edges of a
bidirectional transformation, leaving in the middle a pure bijective
transformation.  
Finally,
we leverage this normalization theorem to {\em synthesize} QRE lenses from a pair
of QREs and example input-output pairs, reusing earlier work on
synthesizing plain bijective lenses.
We have implemented QREs and QRE lens synthesis 
in the bidirectional programming language Boomerang.
We evaluate the effectiveness of our approach by using it to
synthesize QRE lenses between various real-world data formats in
the Optician benchmark suite.
